Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-10-23 decorrelation/decorrelation.tex
including a bit more detail, feel free to revert this commit

One particular way of realizing randomized decoupling is Frame Randomization\cite{Wallman2016, Ware2018}. Frame Randomization can be used to average over noise by twirling error over an appropriately selected unitary 2-design, by inserting an additional precompilation step in the execution of quantum circuits.\cite{roy2009unitary} This has been demonstrated in existing systems\cite{Ware2018}, but can require intensive use of classical processors and waveform memory, and becomes more difficult when the gates being used are not in the Clifford group\needcite, as is the case in various quantum computing architectures.\needcite

Another approach is to consider constructing a family of unitaries that approximate a given gate, and averaging over them. In \cite{Campbell2017}, Campbell considers coherent error arising from unitary synthesis approximations algorithms like the Solovay-Kitaev Algorithm \cite{1612.01011}. He presents an algorithm that, given an oracle for generating unitaries close to a target, produces controls that are quadratically better in diamond norm than any of the individual controls generated by the oracle. In that work, the primary emphasis was on resource counting, and diamond-norm error-bounds at level of unitary synthesis. In this paper, we explore a different but related problem: synthesizing physical controls and making them robust to drift.

We propose to inject additional decorrelating randomness into the system during physical gate synthesis through the use of \emph{robustly balanced controls} (RBCs). The task of generating RBCs will fall to optimal control.

@@

\section{Theory}
 Consider a family of control solutions, $c_i(t)$, where each member of the family approximates the same target gate with slightly different errors. That is, for a target gate $U_T$, the family consists of gates $U_i = \mathcal{T}e^{-i\int c_i(t)H(t, \vec{\delta})}$, where $\vec{\delta}$ are \textit{quasi-static}\cite{Ball2016} parameters of the Hamiltonian. Such a family of unitary approximations is called a \emph{balanced control} if there is a weight vector $\vec{\omega}$ that satisfies the following for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

where $\vec{\omega}$ must be a proper probability distribution and $DPN[\alpha](\rho)$ is a depolarizing noise channel, with strength $\alpha$. Such a channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one.

This is equivalent to the condition given by Campbell\cite{Campbell2017}, that there exist a weight such that

\begin{equation}\label{eq:campbell-condition}
\sum \omega_j H_j(t, \vec{\delta})|_{\delta=\vec{0}} = 0
\end{equation}
However, we will be interested in a stronger condition. A balanced control is said to be \textit{robust to order $\ell$} if a generalized form of Equation \ref{eq:campbell-condition} holds, namely if for all $j \leq \ell$:
\begin{equation}\label{eq:RBC}
D_\ell = \sum_j \omega_j \frac{\partial^{\ell}}{\partial\delta_{i_1}...\partial\delta{i_\ell}}H_j(t,\vec{\delta})|_{\delta=\vec{0}}  = 0
\end{equation}

These conditions imply that an $\ell^{th}$ order RBC ($\ell$RBC) is insensitive to the $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, write the error on each gate in the RBC as $U_j(\vec{\delta}) = exp(-i(\tilde{H}_j(\vec{0}) + \frac{\partial}{\partial\delta_i}\tilde{H}_j(d\delta_i) +  \frac{1}{2}\frac{\partial^2}{\partial\delta_i\partial\delta_k} \tilde{H}_j(d\delta_i d\delta_k) + ...)U_T$, where $\tilde{H}(\delta)$ generates the error $U_j(\vec{\delta})U_T^{\dagger}$ over variations in $\vec{\delta}$. In this way, an $\ell$RBC satisfies Equation \ref{eq:1} to $\ell^{th}$ order.

@@

The question still remains -- when does this quadratic program have a solution. Naturally, if the control family is too small a solution won't exist. In addition, as in \cite{Campbell2017}, we need  $0\in $ Conv$[\{\frac{\partial^{\ell}}{\partial\delta_{i_1}...\partial\delta{i_\ell}}\tilde{H}_j(t,\vec{\delta})|_{\vec{\delta}=0}\}]$. This can be accomplished by modifying the oracle in Campbell's algorithm to produce robust unitaries.

>>>>

One particular way of realizing randomized decoupling is Frame Randomization\cite{Wallman2016, Ware2018}. Frame Randomization can be used to average over noise by twirling error over an appropriately selected unitary 2-design, by inserting an additional precompilation step in the execution of quantum circuits.\cite{roy2009unitary} This has been demonstrated in existing systems\cite{Ware2018}, but can require intensive use of classical processors and waveform memory.

Another way of performing randomized deocupling is through mixing unitaries. In \cite{Campbell2017}, Campbell presents an algorithm that, given an oracle for generating unitaries close to a target, produces controls that are quadratically better in diamond norm than any of the individual controls generated by the oracle. In Campbell's work, he discusses such an oracle could be implemented via unitary synthesis approximations algorithms\cite{1612.01011}, and considers estimating and reducing resources costs that might arise during circuit compilation. In this paper we will we explore a different but related problem: synthesizing physical controls and making them robust to drift. Using optimal control routines as our oracle , we propose to inject additional decorrelating randomness into the system during physical gate synthesis through the use of \emph{robustly balanced controls} (RBCs).

@@

\section{Robustly Balanced Controls}
 Consider the problem of trying to approximate a target unitary operation, $U_T$. Define a control solution $c_i(t)$ by  \begin{equation} 
\mathcal{T}e^{-i\int c_i(t)H(t, \vec{\delta})} = U_i \approx U_T
 \end{equation}
 where $\vec{\delta}$ are meant to be \textit{quasi-static}\cite{Ball2016} parameters of a Hamiltonian, $H(t, \vec{\delta})$. A family of control solutions is called a \emph{balanced control} if there is probability distribution $\vec{\omega}$ that satisfies the following for some small $\alpha$,
\begin{equation}\label{eq:1}
  \sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

where $\vec{\omega}$ must be a proper probability distribution and $DPN[\alpha](\rho)$ is a depolarizing noise channel, with strength $\alpha$. A depolarizing channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one.

This is equivalent to the condition given by Campbell in \cite{Campbell2017}, that there exist a weight such that

\begin{equation}\label{eq:campbell-condition}
\sum \omega_j H_j(t, \vec{\delta})|_{\delta=\vec{0}} = 0
\end{equation}
however we will generalize this condition to define an RBC. A balanced control is said to be robust to order $\ell$ if a generalized form of Equation \ref{eq:campbell-condition} holds, namely if for all $1 \leq j \leq \ell$:
\begin{equation}\label{eq:RBC}
\begin{gathered}
\sum_k\omega_k(\sum_{n=0}^j D^n_k)^j = 0\\
D^n_k = \frac{1}{n!}\frac{\partial^{n}}{\partial\delta_{i_1}...\partial\delta{i_n}}H_k(t,\vec{\delta})|_{\delta=\vec{0}}
\end{gathered}
\end{equation}
Where the first sum is meant to result from Taylor expansion, and therefore will require matrix reshaping. These conditions imply that an $\ell^{th}$ order RBC ($\ell$RBC) is insensitive to the $\ell^{th}$ order in drift in $\vec{\delta}$. To see this, write the error on each gate in the RBC as $U_j(\vec{\delta}) = exp(-i(\tilde{H}_j(\vec{0}) + \frac{\partial}{\partial\delta_i}\tilde{H}_j(d\delta_i) +  \frac{1}{2}\frac{\partial^2}{\partial\delta_i\partial\delta_k} \tilde{H}_j(d\delta_i d\delta_k) + ...)U_T$, where $\tilde{H}(\delta)$ generates the error $U_j(\vec{\delta})U_T^{\dagger}$ over variations in $\vec{\delta}$. Then, by Taylor expanding $U_j(\vec{\delta})U_T^{\dagger}$ one finds that the the first $\ell$ derivatives of an $\ell RBC$ will be zero.

@@

The question still remains -- when does this quadratic program have a solution. Naturally, if the control family is too small a solution won't exist. As in \cite{Campbell2017}, we need  $0\in $ Conv$[\{\frac{\partial^{\ell}}{\partial\delta_{i_1}...\partial\delta{i_\ell}}\tilde{H}_j(t,\vec{\delta})|_{\vec{\delta}=0}\}]$. This can be accomplished by modifying the oracle in Campbell's algorithm to produce robust controls, in particular by updating the GRAPE cost function to include derivative information, one could produce controls with derivatives of order $\epsilon$. Finally one would like to ensure that the control family produced is not too large - that is given the family of controls, we would like to regularize the cost function to enforce sparsity. In this case, lasso regularization is insufficient as we already constrain the one norm of the vector we optimize over to be unity. However, the problem of enforcing sparsity in such situations has been considered in \cite{NIPS2012_4504} and remains a convex problem.
